---
title: "Enggano Impact Questionnaires Results"
author: 
  - id: first
    number: 1
    name:
      given: Gede Primahadi Wijaya
      family: Rajeg
    orcid: 0000-0002-2047-8621
    affiliations:
      - ref: unud
      - ref: cirhss
  - id: second
    number: 2
    name:
      given: Charlotte
      family: Hemmings
    orcid: 0000-0002-3076-5544
    affiliations:
      - ref: oxford
  - id: third
    number: 3
    name:
      given: Engga Zakaria
      family: Sangian
    orcid: 0009-0000-8802-6819
    affiliations:
      - ref: cirhss
  - id: fourth
    number: 4
    name:
      given: Dendi
      family: Wijaya
    orcid: 0000-0002-8767-9364
    affiliations:
      - ref: brin
  - id: fifth
    number: 5
    name:
      given: I Wayan
      family: Arka
    orcid: 0000-0002-2819-6186
    affiliations:
      - ref: anu
      - ref: cirhss
  - id: sixth
    number: 6
    name:
      given: Mary
      family: Dalrymple
    orcid: 0000-0001-5723-5722
    affiliations:
      - ref: oxford
affiliations:
  - id: unud
    name: Universitas Udayana
  - id: cirhss
    name: Centre for Interdisciplinary Research on the Humanities and Social Sciences, Univ. Udayana
  - id: oxford
    name: University of Oxford
  - id: brin
    name: National Research and Innovation Agency
  - id: anu
    name: Australian National University
format:
  docx:
    toc: true
    toc-depth: 3
    warning: false
    message: false
    fig-width: 6
    fig-asp: 0.618
    fig-dpi: 300
    fig-align: "center"
    number-sections: true
    echo: false
  pdf:
    message: false
    toc: true
    warning: false
    fig-width: 6
    fig-asp: 0.618
    fig-dpi: 300
    fig-align: "center"
    number-sections: true
    fig-pos: "H"
    echo: false
    geometry:
      - top=1in
      - left=1in
      - right=1in
      - bottom=1in
    
  html:
    code-fold: true
    code-block-border-left: "#31BAE9"
    code-block-bg: "#F8F9F9"
    message: false
    toc: true
    number-sections: true
    warning: false
    smooth-scroll: true
    fig-width: 6
    fig-asp: 0.618
    fig-dpi: 300
    fig-align: "center"
editor: visual
date: 2025-06-23
date-modified: now
---

# Preface

This is a [Quarto](https://quarto.org/ "Quarto Webpage"){target="_blank"}-generated report for the results of the Enggano Impact Questionnaires, which were collected in the Enggano Island by Engga Zakaria Sangian between June 6 and June 13, 2025.

```{r setup}
#| message: false
#| warning: false

# a setup chunk
library(likert)
library(grid)
source("pre-processing.R") # all loading the tidyverse
```

## Metadata Question No. 4

It asks whether respondents are (i) teachers (in elementary school \[*SD*\], junior high school \[*SMP*\], or senior high school \[*SMA*\], or (ii) non-teachers (*BUKAN GURU*) (see @fig-q1). Numbers in brackets are raw counts.

```{r}
#| label: fig-q1
#| fig-cap: "Distribution of Teachers vs. Non-Teachers for Metadata Question no. 4"

q1 <- quiz |> 
  select(mengajar = `4. Mengajar di tingkat?`) |> 
  mutate(mengajar1 = if_else(str_detect(mengajar, "^(BUKAN|^Pegawai)"), "BUKAN GURU", mengajar)) |> 
  mutate(mengajar1 = str_split(mengajar1, "\\, ")) |> 
  unnest_longer(mengajar1)

q1_count <- q1 |> 
  mutate(mengajar2 = "non-teacher",
         mengajar2 = replace(mengajar2, mengajar1 == "SD", "elementary school teacher"),
         mengajar2 = replace(mengajar2, mengajar1 == "SMP", "junior high school teacher"),
         mengajar2 = replace(mengajar2, mengajar1 == "SMA", "senior high school teacher")) |> 
  mutate(mengajar2 = factor(mengajar2, levels = c("elementary school teacher",
                                                  "junior high school teacher",
                                                  "senior high school teacher",
                                                  "non-teacher"))) |> 
  count(mengajar2) |> 
  mutate(perc = round(n/sum(n)*100, 1))

q1_count |> 
  ggplot(aes(x = mengajar2, y = n)) + 
  geom_col() +
  labs(x = "Is teacher? At which level?",
       y = "frequency") +
  geom_text(aes(label = str_c(perc, "% (", n, ")", sep = "")),
            colour = "white",
            vjust = 1.5,
            size = 3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

Note that there is one respondent who indicates that (s)he teaches at both elementary and junior high schools. For this respondent, h(is/er) responses are counted twice in order to represent each level of school (s)he teaches. This is the reason why the sum of the raw counts in @fig-q1 is `{r} sum(q1_count$n)` instead of `{r} nrow(quiz)`, which is the number of people asked for answering the questionnaires.

# Questions about Textbooks

## Main Question no. 1

This question asks whether the Enggano language has ever been taught in class before the publication of the textbooks for Junior High and Elementary School (@fig-q5)

```{r}
#| label: fig-q5
#| fig-cap: "Responses for Question no. 1"
#| fig-asp: 0.75

q5 <- quiz_chr |> 
  filter(str_detect(question, "^5")) |> 
  count(mengajar2, mengajar3, response) |> 
  mutate(response = if_else(response == "Ya", "Yes", "No"),
         response = factor(response, levels = c("Yes", "No"))) |> 
  mutate(perc = round(n/sum(n)*100, 1))

ggplot(q5, aes(x = mengajar2, y = n, fill = response)) +
  geom_col(position = "dodge") +
  facet_wrap(~mengajar3, scales = "free_x") +
  geom_text(aes(label = str_c(perc, "%\n(", n, ")", sep = "")),
            position = position_dodge(.9),
            vjust = -.25,
            size = 3) +
  theme_bw() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 8)) +
  scale_y_continuous(limits = c(0, 14), breaks = c(2, 6, 10, 14), minor_breaks = NULL) +
  labs(x = NULL,
       y = "frequency",
       caption = "Has Enggano been taught before the production of the textbooks?")
```

## Main Question no. 2

The question asked whether the Enggano language subject is compulsory in the curriculum today (@fig-q6) This question could also be interpreted by the respondents as Enggano should be a formally/regulated compulsory subject. There is an information that the local/district level government instructs for the teaching of the language at school but that has not been formalised as a written regulation.

```{r}
#| label: fig-q6
#| fig-cap: "Responses for Question no. 2"
#| fig-asp: 0.75

q6 <- quiz_chr |> 
  filter(str_detect(question, "^6")) |> 
  count(mengajar2, mengajar3, response) |> 
  mutate(response = if_else(response == "Ya", "Yes", "No"),
         response = factor(response, levels = c("Yes", "No"))) |> 
  mutate(perc = round(n/sum(n)*100, 1))

ggplot(q6, aes(x = mengajar2, y = n, fill = response)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = str_c(perc, "% (", n, ")", sep = "")),
            position = position_dodge(.9),
            vjust = -.25,
            size = 3) +
  theme_bw() +
  facet_wrap(~mengajar3, scales = "free_x") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 8)) +
  scale_y_continuous(limits = c(0, 12), breaks = c(2, 6, 10, 14), 
                     minor_breaks = NULL) +
  labs(x = NULL,
       y = "frequency",
       caption = "Is Enggano a compulsory subject in the curriculum?")

```

## Main Question no. 3 {#sec-q3}

We ask whether the textbooks for each school level (Junior High and Elementary levels) have ever been used in class since we released them in January 2025. We ask respondents to be honest if the books have never been used (see @fig-q7)

```{r}
#| label: fig-q7
#| fig-cap: "Responses for Question no. 3"
#| fig-asp: 0.75

q7 <- quiz_chr |> 
  filter(str_detect(question, "^7")) |> 
  count(mengajar2, mengajar3, response) |> 
  mutate(response = if_else(response == "ya, sudah pernah", "Yes", "Never"),
         response = factor(response, levels = c("Yes", "Never"))) |> 
  mutate(perc = round(n/sum(n)*100, 1))

ggplot(q7, aes(x = mengajar2, y = n, fill = response)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = str_c(perc, "% (", n, ")", sep = "")),
            position = position_dodge(.9),
            vjust = -.25,
            size = 3) +
  theme_bw() +
  facet_wrap(~mengajar3, scales = "free_x") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 8)) +
  scale_y_continuous(limits = c(0, 12), breaks = c(2, 6, 10, 14), 
                     minor_breaks = NULL) +
  labs(x = NULL,
       y = "frequency",
       caption = "Have the produced textbooks been used in classroom teaching?")
```

## Main Question no. 4 {#sec-q8}

When in @sec-q3 the respondents answered "**Never (used the textbook in class)**", or when they are not teachers, but have seen the contents of the book, we asked them to give an impression of how difficult it seems to use the primary and/or junior high school book, on the 5-point Likert scale. The point 1 means 'very difficult' up to point 5 'very easy'. We can interpret the point 3 as intermediate/neutral impression. Below, we also provide an example of interpreting this Likert-scale bar plots (see @fig-q8) generated using the R package `likert` and `grid`, which can be useful for other same kind of plots.

```{r}
#| fig.width: 9
#| label: fig-q8
#| fig-cap: "Rating results for Question no. 4"

q8 <- quiz_dbl |> 
  filter(str_detect(question, "^8")) |> 
  mutate(response = factor(response),
         response = factor(response, levels = c("1", levels(response)))) |> 
  select(mengajar2, 
         mengajar3, 
         `How difficult it seems to use the textbooks in class?` = response)

q8_count <- q8 |> 
  group_by(mengajar2) |> 
  filter(!is.na(`How difficult it seems to use the textbooks in class?`)) |> 
  count(`How difficult it seems to use the textbooks in class?`) |> 
  mutate(perc = round(n/sum(n)*100, 1)) |> 
  rename(quest = `How difficult it seems to use the textbooks in class?`)

q8_resp <- as.data.frame(q8[,3])
q8_resp_lik <- likert::likert(q8_resp, grouping = q8$mengajar2, nlevels = )
plot(q8_resp_lik, include.histogram = TRUE, include.center = TRUE,
     plot.percents = TRUE)
```

The right panel shows the number of missing (non-answered) responses (maroon) and of completed responses. The left panel shows the distribution of the Likert responses by respondents. Here is an example to interpret the left graph. Take the `non-teacher` response. The `{r} round(q8_resp_lik$results[q8_resp_lik$results$Group == "non-teacher", "2"])`% on the left indicates the cumulative proportion of respondents choosing point 2. So, for this group of respondents, the textbooks appear to be difficult to use, because there are only `{r} round(q8_resp_lik$results[q8_resp_lik$results$Group == "non-teacher", "5"])`% responses capturing point 5 ('very easy'). On the contrary, for the `elementary school teacher`, the impression is reverse whereby cumulatively (up to `{r} round(sum(q8_resp_lik$results[q8_resp_lik$results$Group == "elementary school teacher", c("4", "5")]))`%), this group perceives the textbooks to be relatively easy to use (because selecting points 4 and 5). @tbl-q8 provides the raw counts for @fig-q8. The values inside the `percentage` column is calculated by the `group` column.

```{r}
#| tbl-cap: "Counts of the rating in Question 4"
#| label: tbl-q8
#| warning: false
#| message: false

q8_count |> 
  rename(group = mengajar2,
         rating_scale = quest,
         no_of_voter = n,
         percentage = perc) |> 
  knitr::kable()
```

## Main Question no. 5

In relation to the previous question (i.e., Main Question no. 4 in @sec-q8), are there any other suggestions/comments you would like to make regarding the primary and junior high school textbooks? The responses are shown in @tbl-q8a .

```{r}
#| message: false
#| warning: false
#| tbl-colwidths: [20, 40, 40]
#| tbl-cap: "Responses for Main Question no. 5"
#| label: tbl-q8a


q8a <- read_tsv("q8-trans-to-edit.tsv")
q8a |> 
  mutate(mengajar3 = factor(mengajar3, levels = c("teachers", "non-teachers")),
         mengajar2 = factor(mengajar2, levels = c("elementary school teacher",
                                                  "junior high school teacher",
                                                  "senior high school teacher",
                                                  "non-teacher"))) |> 
  arrange(mengajar3, mengajar2) |> 
  select(`Respondent` = mengajar2, Indonesian = response, English = resp_trans_edit) |> 
  knitr::kable()
```

## Main Question no. 6 {#sec-q6}

When in @sec-q3 the respondents answered "Yes" that they have used the textbooks in their classes, we asked them to rate (on the 5-point scale) how difficult it would be for the teacher to teach the material in the textbooks. The interpretation is the same, point 1 is very difficult and point 5 is very easy (see @fig-q9).

```{r}
#| fig.width: 9
#| label: fig-q9
#| fig-cap: "Rating results for Question no. 6"

q9 <- quiz_dbl |> 
  filter(str_detect(question, "^9")) |> 
  mutate(response = factor(response),
         response = factor(response, levels = c("1", levels(response)))) |> 
  select(mengajar2, 
         mengajar3, 
         `How difficult it seems to use the textbooks in class?` = response)

q9_count <- q9 |> 
  group_by(mengajar2) |> 
  filter(!is.na(`How difficult it seems to use the textbooks in class?`)) |> 
  count(`How difficult it seems to use the textbooks in class?`) |> 
  mutate(perc = round(n/sum(n)*100, 1)) |> 
  rename(quest = `How difficult it seems to use the textbooks in class?`)

q9_resp <- as.data.frame(q9[,3])
q9_resp_lik <- likert::likert(q9_resp, grouping = q9$mengajar2, nlevels = )
plot(q9_resp_lik, include.histogram = TRUE, include.center = TRUE,
     plot.percents = TRUE)
```

Note that the `{r} pull(filter(q9_count, mengajar2 == "non-teacher"), perc)` % for the rather neutral scale (point 3) in the `non-teacher` group is represented by just a single respondent. The raw counts are shown in @tbl-q9.

```{r}
#| tbl-cap: "Counts of the rating in Question 6"
#| label: tbl-q9
#| warning: false
#| message: false

q9_count |> 
  rename(group = mengajar2,
         rating_scale = quest,
         no_of_voter = n,
         percentage = perc) |> 
  knitr::kable()
```

## Main Question no. 7 {#sec-q10}

This is a leading, Boolean (yes-no) question about any difficult parts of the textbooks for the teachers to teach. Options regarding which parts are provided in @sec-q11.

```{r}
#| label: fig-q10
#| fig-cap: "Responses for Question no. 7"
#| fig-asp: 0.75

q10 <- quiz_chr |> 
  filter(str_detect(question, "^10")) |> 
  count(mengajar2, mengajar3, response) |> 
  mutate(response = if_else(response == "Ya, ada bagian yang sulit", "Yes, there are certain difficult parts", "None"),
         response = factor(response, levels = c("Yes, there are certain difficult parts", "None"))) |> 
  mutate(perc = round(n/sum(n)*100, 1))

q10 |> 
  ggplot(aes(x = mengajar2, y = n, fill = response)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = str_c(perc, "% (", n, ")", sep = "")),
            position = position_dodge(.9),
            vjust = -0.35,
            size = 3) +
  theme_bw() +
  facet_wrap(~mengajar3, scales = "free_x") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 8)) +
  scale_y_continuous(limits = c(0, 18), breaks = c(2, 6, 10, 14, 18), 
                     minor_breaks = NULL) +
  labs(x = NULL,
       y = "frequency",
       caption = "Are there certain parts of the textbooks that teachers find difficult to teach?")
```

## Main Question no. 8 {#sec-q11}

This question is a follow-up for the question in @sec-q10. It asks respondents to multi-choose from a pre-determined list, with an option to add "others", which parts they find difficult to teach (@fig-q11).

```{r}
#| fig-asp: 0.85
#| label: fig-q11
#| fig-cap: "Responses for Question no. 8"

q11 <- quiz_chr |> 
  filter(str_detect(question, "^11")) |> 
  mutate(response = str_split(response, ", ")) |> 
  unnest_longer(response) |> 
  count(mengajar2, mengajar3, response)
q11_resp <- tribble(~main_cat, ~response, ~english,
                    TRUE, "berbicara/pelafalan bahasa Enggano", "speaking/pronunciation",
                    TRUE, "kosakata bahasa Enggano", "vocabulary",
                    TRUE, "membaca bahasa Enggano", "reading",
                    TRUE, "menulis bahasa Enggano", "writing/orthography",
                    TRUE,"tatabahasa bahasa Enggano", "grammar",
                    FALSE, "Untuk penulisan tanda baca dan pembaca memang masih harus di singkronkan antra para pengajar agar mudah memaparkan dan mengajarkan pada para siswa", "We still need to syncronise the understanding of the readers with the orthography/spelling. That way, it would be easier between teachers to explain and teach the orthography to the students.")
q11 <- left_join(q11, q11_resp, by = join_by("response")) |> 
  mutate(main_cat = replace(main_cat, is.na(main_cat), TRUE)) |> 
  filter(main_cat) |> 
  mutate(perc = round(n/sum(n)*100, 1))

q11 |> 
  ggplot(aes(x = mengajar2, y = n, fill = english)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = str_c(perc, "% (", n, ")", sep = "")),
            position = position_dodge(.9),
            hjust = -0.15,
            size = 2.25, fontface = "bold") +
  theme_bw() +
  facet_wrap(~mengajar3, scales = "free_y") +
  theme(legend.position = "top",
        legend.text = element_text(size = 8),
        axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 18), breaks = c(2, 6, 10, 14, 18), 
                     minor_breaks = NULL) +
  coord_flip() +
  labs(x = NULL,
       y = "frequency",
       fill = "response",
       caption = "Which parts of the textbooks are difficult to teach?")
```

There is one respondent choosing "others" and providing a comment regarding orthography as follows:

> Indonesian: "*Untuk penulisan tanda baca dan pembaca memang masih harus di singkronkan antra para pengajar agar mudah memaparkan dan mengajarkan pada para siswa*"
>
> English: "We still need to syncronise the understanding of the readers with the orthography/spelling. That way, it would be easier between teachers to explain and teach the orthography to the students."

## Main Question no. 9

If in the Main Question no. 7 (@sec-q10) you answered ‘Yes’ that there are certain difficult parts of the textbook to teach, please provide comments regarding solutions that you can think of to reduce these difficulties.

```{r}
#| message: false
#| warning: false
#| tbl-colwidths: [20, 40, 40]
#| tbl-cap: "Responses for Main Question no. 9"
#| label: tbl-q12


q12 <- read_tsv("q12-trans-to-edit.tsv")
q12 |> 
  mutate(mengajar3 = factor(mengajar3, levels = c("teachers", "non-teachers")),
         mengajar2 = factor(mengajar2, levels = c("elementary school teacher",
                                                  "junior high school teacher",
                                                  "senior high school teacher",
                                                  "non-teacher"))) |> 
  arrange(mengajar3, mengajar2) |> 
  select(`Respondent` = mengajar2, Indonesian = response, English = resp_trans_edit) |> 
  knitr::kable()
```

# Questions for Digital and Print Dictionaries

## Main Question no. 10 {#sec-q13}

This question asks whether the Digital Enggano Language Dictionary and the printed Enggano language dictionary have ever been used in the classroom. We invite the respondents to be honest if they have not used them. The results are shown in @fig-q13.

```{r}
#| fig-asp: 0.85
#| fig-width: 7
#| label: fig-q13
#| fig-cap: "Responses for Question no. 10"

q13 <- quiz_chr |> 
  filter(str_detect(question, "^13\\.")) |> 
  count(mengajar2, mengajar3, response) |> 
  mutate(perc = round(n/sum(n) * 100, 1)) |> 
  mutate(response_english = response) |> 
  mutate(response_english = replace(response_english,
                                    response_english == "Belum pernah kedua-duanya",
                                    "never for both"),
         response_english = replace(response_english,
                                    response_english == "Ya, sudah pernah Kamus Cetak saja",
                                    "yes, print only"),
         response_english = replace(response_english,
                                    response_english == "Ya, sudah pernah Kamus Digital saja",
                                    "yes, digital only"))

q13 |> 
  ggplot(aes(x = mengajar2, y = n, fill = response_english)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = str_c(perc, "% (", n, ")", sep = "")),
            position = position_dodge(.9),
            hjust = ifelse(q13$perc >= 20, 1.5, -0.15),
            size = 2.25, fontface = "bold") +
  theme_bw() +
  coord_flip() +
  facet_wrap(~mengajar3, scales = "free_y") +
  theme(legend.position = "top",
        legend.text = element_text(size = 8),
        axis.text.y = element_text(size = 8)) +
  labs(x = NULL,
       y = "frequency",
       fill = "response",
       caption = "Have the two versions of the dictionary been used in class?")
```

## Main Question no. 11

When in @sec-q13 the respondents answered “never” used both the printed and digital/online dictionaries, but have seen the contents, we asked them to give an impression about how difficult it seems to use each dictionary. The interpretation is the same, point 1 is very difficult and point 5 is very easy (see @fig-q14).

```{r}
#| fig-asp: 0.75
#| fig-width: 9
#| label: fig-q14
#| fig-cap: "Responses for Question no. 11"

q14 <- quiz_dbl |> 
  filter(str_detect(question, "^14")) |> 
  mutate(response = factor(response),
         response = factor(response, levels = c("1", levels(response)))) |> 
  select(mengajar2, 
         mengajar3, 
         `How difficult it seems to use the two dictionaries?` = response)

q14_count <- q14 |> 
  group_by(mengajar2) |> 
  filter(!is.na(`How difficult it seems to use the two dictionaries?`)) |> 
  count(`How difficult it seems to use the two dictionaries?`) |> 
  mutate(perc = round(n/sum(n)*100, 1)) |> 
  rename(quest = `How difficult it seems to use the two dictionaries?`)

q14_resp <- as.data.frame(q14[,3])
q14_resp_lik <- likert::likert(q14_resp, grouping = q14$mengajar2, nlevels = )
plot(q14_resp_lik, include.histogram = TRUE, include.center = TRUE,
     plot.percents = TRUE)

```

The raw count for @fig-q14 is shown in @tbl-q14. Note that the `senior high school teacher` responses are only from `{r} pull(tally(filter(q14_count, mengajar2 == "senior high school teacher"), n), n)` respondents in total.

```{r}
#| tbl-cap: "Counts of the rating in Question 11"
#| label: tbl-q14
#| warning: false
#| message: false

q14_count |> 
  rename(group = mengajar2,
         rating_scale = quest,
         no_of_voter = n,
         percentage = perc) |> 
  knitr::kable()
```

## Main Question no. 12 {#sec-q15}

In this question, we asked how difficult it is for teachers to use the **Digital** dictionary in the classroom via a 5-point likert scale. This question is relevant especially if respondents answered "yes" in @sec-q13 (i.e., yes, having used the Digital dictionary only or together with the Print one). The interpretation is the same, point 1 is very difficult and point 5 is very easy (see @fig-q15). The raw counts are shown in @tbl-q15.

```{r}
#| fig-asp: 0.75
#| fig-width: 9
#| label: fig-q15
#| fig-cap: "Responses for Question no. 12"

q15 <- quiz_dbl |> 
  filter(str_detect(question, "^15")) |> 
  mutate(response = factor(response),
         response = factor(response, levels = c("1", levels(response)))) |> 
  select(mengajar2, 
         mengajar3, 
         `How difficult it is to use the Digital dictionary?` = response)

q15_count <- q15 |> 
  group_by(mengajar2) |> 
  filter(!is.na(`How difficult it is to use the Digital dictionary?`)) |> 
  count(`How difficult it is to use the Digital dictionary?`) |> 
  mutate(perc = round(n/sum(n)*100, 1)) |> 
  rename(quest = `How difficult it is to use the Digital dictionary?`)

q15_resp <- as.data.frame(q15[,3])
q15_resp_lik <- likert::likert(q15_resp, grouping = q15$mengajar2, nlevels = )
plot(q15_resp_lik, include.histogram = TRUE, include.center = TRUE,
     plot.percents = TRUE)
```

```{r}
#| tbl-cap: "Counts of the rating in Question 12"
#| label: tbl-q15
#| warning: false
#| message: false

q15_count |> 
  rename(group = mengajar2,
         rating_scale = quest,
         no_of_voter = n,
         percentage = perc) |> 
  knitr::kable()
```

## Main Question no. 13 {#sec-q16}

In this question, we asked how difficult it is for teachers to use the **Print** dictionary in the classroom via a 5-point likert scale. This question is relevant especially if respondents answered "yes" in @sec-q13 (i.e., yes, having used the Print dictionary only or together with the Print one). The interpretation is the same, point 1 is very difficult and point 5 is very easy (see @fig-q16). The raw counts are shown in @tbl-q16.

```{r}
#| fig-asp: 0.75
#| fig-width: 9
#| label: fig-q16
#| fig-cap: "Responses for Question no. 13"

q16 <- quiz_dbl |> 
  filter(str_detect(question, "^16")) |> 
  mutate(response = factor(response),
         response = factor(response, levels = c("1", levels(response)))) |> 
  select(mengajar2, 
         mengajar3, 
         `How difficult it is to use the Print dictionary?` = response)

q16_count <- q16 |> 
  group_by(mengajar2) |> 
  filter(!is.na(`How difficult it is to use the Print dictionary?`)) |> 
  count(`How difficult it is to use the Print dictionary?`) |> 
  mutate(perc = round(n/sum(n)*100, 1)) |> 
  rename(quest = `How difficult it is to use the Print dictionary?`)

q16_resp <- as.data.frame(q16[,3])
q16_resp_lik <- likert::likert(q16_resp, grouping = q16$mengajar2, nlevels = )
plot(q16_resp_lik, include.histogram = TRUE, include.center = TRUE,
     plot.percents = TRUE)
```

```{r}
#| tbl-cap: "Counts of the rating in Question 13"
#| label: tbl-q16
#| warning: false
#| message: false

q16_count |> 
  rename(group = mengajar2,
         rating_scale = quest,
         no_of_voter = n,
         percentage = perc) |> 
  knitr::kable()
```

## Main Question no. 14 {#sec-q17}

As the question in @sec-q11, in this question we ask whether there are any aspects/parts of the **Digital** dictionary that users find difficult (esp. if they have seen and used it). One participant can choose more than one option. The results are shown in @fig-q17.

```{r}
#| fig-width: 9
#| label: fig-q17
#| fig-cap: "Responses for Question no. 14"

q17 <- quiz_chr |> 
  filter(str_detect(question, "^17")) |> 
  mutate(response = str_replace_all(response, 
                                    "Baru tau ketika Engga, datang ke Enggano.",
                                    "Baru tau ketika Engga datang ke Enggano.")) |> 
   mutate(response = str_split(response, ", ")) |> 
  unnest_longer(response) |> 
  count(mengajar2, mengajar3, response)
q17_resp <- tribble(~main_cat, ~response, ~english,
                    TRUE, "Pemilihan bahasa awal untuk memulai pencarian kata", "choosing the search language",
                    TRUE, "Pengetikan huruf dengan simbol-simbol khusus", "typing special characters/orthography",
                    TRUE, "Pencarian kata", "word search",
                    TRUE, "Pemahaman isi/informasi dari hasil pencarian kata", "understanding output of the search result",
                    FALSE, "Baru tau ketika Engga datang ke Enggano.", "Just knew it when Engga came to Enggano.")
q17 <- left_join(q17, q17_resp, by = join_by("response")) |> 
  mutate(main_cat = replace(main_cat, is.na(main_cat), TRUE)) |> 
  filter(main_cat) |> 
  mutate(perc = round(n/sum(n)*100, 1))

q17 |> 
  ggplot(aes(x = mengajar2, y = n, fill = english)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = str_c(perc, "% (", n, ")", sep = "")),
            position = position_dodge(.9),
            hjust = -0.15,
            size = 2.25, fontface = "bold") +
  theme_bw() +
  facet_wrap(~mengajar3, scales = "free_y") +
  theme(legend.text = element_text(size = 8),
        axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 18), breaks = c(2, 6, 10, 14, 18), 
                     minor_breaks = NULL) +
  coord_flip() +
  labs(x = NULL,
       y = "frequency",
       fill = "response",
       caption = "Which parts of the Digital dictionary are difficult?")
```

One respondent chose "others", and saying that (s)he just knew about the dictionary because Engga came again to Enggano in June.

## Main Question no. 15 {#sec-q18}

Following on the question in @sec-q17, in this question we ask whether there are any aspects/parts of the **Print** dictionary that users find difficult (esp. if they have seen and used it). One participant can choose more than one option. The results are shown in @fig-q18.

```{r}
#| fig-width: 9
#| label: fig-q18
#| fig-cap: "Responses for Question no. 15"

q18 <- quiz_chr |> 
  filter(str_detect(question, "^18")) |> 
   mutate(response = str_split(response, ", ")) |> 
  unnest_longer(response) |> 
  count(mengajar2, mengajar3, response)
q18_resp <- tribble(~main_cat, ~response, ~english,
                    TRUE, "Pelafalan huruf-huruf dengan simbol khusus", "pronouncing words with special orthography",
                    TRUE, "Penelusuran kata di bahasa Enggano", "word search in Enggano",
                    TRUE, "Penelusuran kata di bahasa Indonesia", "word finder via Indonesian",
                    TRUE, "Struktur informasi yang ditampilkan untuk suatu kata", "information inside a word entry",
                    FALSE, "Perbanyak kamus cetak", "Please make more copies of the dictionary")
q18 <- left_join(q18, q18_resp, by = join_by("response")) |> 
  mutate(main_cat = replace(main_cat, is.na(main_cat), TRUE)) |> 
  filter(main_cat) |> 
  mutate(perc = round(n/sum(n)*100, 1))

q18 |> 
  ggplot(aes(x = mengajar2, y = n, fill = english)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = str_c(perc, "% (", n, ")", sep = "")),
            position = position_dodge(.9),
            hjust = -0.15,
            size = 2.25, fontface = "bold") +
  theme_bw() +
  facet_wrap(~mengajar3, scales = "free_y") +
  theme(legend.text = element_text(size = 8),
        axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 18), breaks = c(2, 6, 10, 14, 18), 
                     minor_breaks = NULL) +
  coord_flip() +
  labs(x = NULL,
       y = "frequency",
       fill = "response",
       caption = "Which parts of the Print dictionary are difficult?")
```

One respondent requested to print more copies of the dictionary.

# Future aspiration for maximising the textbooks and dictionaries

## Main Question no. 16 {#sec-q19}

### Context

In January 2025, during the delivery of research results (textbooks and printed dictionaries), the Enggano language research team did not have the opportunity to conduct workshops/training on the use of textbooks and further introduction to dictionaries (due to weather + transportation constraints and return tickets that had already been booked). We will prepare a video tutorial related to the material in question (here is [the one for the Digital dictionary](https://youtube.com/playlist?list=PLN_hyg71EM-fyuz8v_jw9xmp9krG4ijHw&si=85LzQC9uRqhoisjS)). We would like to get input regarding what needs to be emphasised in the tutorial through the following questions.

### The question

To maximise the use/understanding of the materials in the **Textbooks** and the **Digital & Print** dictionaries in the classroom, what would teachers/communities most like to learn more through tutorials from the Enggano language research team? The respondents can choose more than one option.

```{r}
#| fig-width: 9
#| label: fig-q19
#| fig-cap: "Responses for Question no. 16"

q19 <- quiz_chr |> 
  filter(str_detect(question, "^19")) |> 
   mutate(response = str_split(response, ", ")) |> 
  unnest_longer(response) |> 
  count(mengajar2, mengajar3, response)
q19_resp <- tribble(~main_cat, ~response, ~english,
                    TRUE, "Tata tulis/Ejaan/Ortografi Bahasa Enggano", "Enggano orthography",
                    TRUE, "Praktik penggunaan buku ajar yang disusun dalam “Rencana Pelaksanaan Pembelajaran” (RPP)", "practice in using the Textbooks laid out in a “Lesson Plan”",
                    TRUE, "Menelusuri kata dalam kamus", "searching words in the Dictionaries",
                    TRUE, "Tatabahasa Bahasa Enggano", "Enggano grammar")
q19 <- left_join(q19, q19_resp, by = join_by("response")) |> 
  mutate(main_cat = replace(main_cat, is.na(main_cat), TRUE)) |> 
  filter(main_cat) |> 
  mutate(perc = round(n/sum(n)*100, 1))

q19 |> 
  ggplot(aes(x = mengajar2, y = n, fill = english)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = str_c(perc, "% (", n, ")", sep = "")),
            position = position_dodge(.9),
            hjust = -0.15,
            size = 2.25, fontface = "bold") +
  theme_bw() +
  facet_wrap(~mengajar3, scales = "free_y") +
  theme(legend.text = element_text(size = 8),
        axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 18), breaks = c(2, 6, 10, 14, 18), 
                     minor_breaks = NULL) +
  coord_flip() +
  labs(x = NULL,
       y = "frequency",
       fill = "response",
       caption = "Aspects to be learnt further by the community via tutorials")
```

## Other open comments

```{r}
q20 <- quiz_chr |> 
  filter(str_detect(question, "^20"))
q20_na <- filter(q20, is.na(response))
```

The question: "Is there anything else that the community needs regarding textbooks and dictionaries? Please write it in the blank space below".

```{r}
#| message: false
#| warning: false
#| tbl-colwidths: [20, 40, 40]
#| tbl-cap: "Responses for the \"Open Comments\""
#| label: tbl-q20


q20 <- read_tsv("q20-trans-to-edit.tsv")
q20 |> 
  mutate(mengajar3 = factor(mengajar3, levels = c("teachers", "non-teachers")),
         mengajar2 = factor(mengajar2, levels = c("elementary school teacher",
                                                  "junior high school teacher",
                                                  "senior high school teacher",
                                                  "non-teacher"))) |> 
  arrange(mengajar3, mengajar2) |> 
  select(`Respondent` = mengajar2, Indonesian = response, English = resp_trans_edit) |> 
  knitr::kable()
```
